---
title: "Demultiplexing Report (BCL Convert)"
author: "`r tryCatch(system('whoami', intern = TRUE)[1], error = function(e) 'unknown')`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  reports_dir: "Reports"
  run_id: ""
output:
  html_document:
    css: styles.css
    keep_md: yes
    number_sections: true
    toc: true
    toc_depth: 3
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 10)
knitr::opts_chunk$set(fig.height = 10)

library("ggplot2")
library("plyr")
library("knitr")

mycat <- function(text) {
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}

make_numeric <- function(x) {
  x <- as.numeric(gsub(pattern = ",", replacement = "", x = x))
  return(x)
}

`%||%` <- function(a, b) if (!is.null(a) && length(a) && nzchar(as.character(a)[1])) a else b

reports_dir <- params$reports_dir
if (!dir.exists(reports_dir)) {
  stop("reports_dir not found: ", reports_dir, call. = FALSE)
}

# run_id is a normal variable — params is read-only in knitr; do not assign to params$run_id
ri <- file.path(reports_dir, "RunInfo.xml")
run_id <- as.character(params$run_id %||% "")
if (run_id == "" && file.exists(ri)) {
  rl <- readLines(ri, warn = FALSE)
  hit <- grep("<Run Id=", rl, value = TRUE)
  if (length(hit)) {
    run_id <- sub('.*<Run Id="([^"]+)".*', "\\1", hit[1])
  }
}

demux <- read.csv(file.path(reports_dir, "Demultiplex_Stats.csv"),
                  check.names = FALSE, stringsAsFactors = FALSE)
qm <- read.csv(file.path(reports_dir, "Quality_Metrics.csv"),
               check.names = FALSE, stringsAsFactors = FALSE)
Samplesheet <- read.csv(file.path(reports_dir, "SampleSheet.csv"),
               check.names = FALSE, stringsAsFactors = FALSE)

mq_mean_col <- "Mean Quality Score (PF)"

qm_agg <- ddply(qm, c("Lane", "SampleID"), function(df) {
  y <- sum(as.numeric(df$Yield))
  yq <- sum(as.numeric(df$YieldQ30))
  mq <- as.numeric(df[[mq_mean_col]])
  wy <- as.numeric(df$Yield)
  data.frame(
    Yield_Mbases = y / 1e6,
    pcnt_Q30 = ifelse(y > 0, yq / y * 100, NA_real_),
    Mean_Q = ifelse(y > 0, sum(wy * mq) / y, NA_real_)
  )
})

demux_key <- paste(demux$Lane, demux$SampleID, sep = ":")
qm_key <- paste(qm_agg$Lane, qm_agg$SampleID, sep = ":")
m <- match(demux_key, qm_key)
demux$Yield_Mbases <- qm_agg$Yield_Mbases[m]
demux$pcnt_Q30 <- qm_agg$pcnt_Q30[m]
demux$Mean_Q <- qm_agg$Mean_Q[m]

barcode_display <- vapply(seq_len(nrow(demux)), function(i) {
  idx <- demux$Index[i]
  if (is.na(idx) || idx == "") return("")
  gsub("-", "+", idx)
}, character(1))

lane_summary_df <- data.frame(
  Lane = demux$Lane,
  Project = Samplesheet[grepl("RunName", Samplesheet$`[Header]`),][,2],
  Sample = demux$SampleID,
  `Barcode sequence` = barcode_display,
  `PF Clusters` = as.numeric(demux[["# Reads"]]),
  `% of thelane` = as.numeric(demux[["% Reads"]]) * 100,
  `% Perfectbarcode` = as.numeric(demux[["% Perfect Index Reads"]]) * 100,
  `% One mismatchbarcode` = as.numeric(demux[["% One Mismatch Index Reads"]]) * 100,
  `Yield (Mbases)` = demux$Yield_Mbases,
  `% PFClusters` = 100,
  `% >= Q30bases` = demux$pcnt_Q30,
  `Mean QualityScore` = demux$Mean_Q,
  check.names = FALSE,
  stringsAsFactors = FALSE
)

lane_summary_numeric_cols <- c(
  "PF Clusters", "% of thelane", "% Perfectbarcode", "% One mismatchbarcode",
  "Yield (Mbases)", "% PFClusters", "% >= Q30bases", "Mean QualityScore"
)

lane_summary_df <- cbind(
  lane_summary_df[, !colnames(lane_summary_df) %in% lane_summary_numeric_cols, drop = FALSE],
  apply(lane_summary_df[, lane_summary_numeric_cols, drop = FALSE], 2, make_numeric)
)

colnames(lane_summary_df) <- gsub(" ", "_", colnames(lane_summary_df), fixed = TRUE)
colnames(lane_summary_df) <- gsub("%", "pcnt", colnames(lane_summary_df), fixed = TRUE)
colnames(lane_summary_df) <- gsub(">=", "greaterthan_equal", colnames(lane_summary_df), fixed = TRUE)
colnames(lane_summary_df) <- gsub("(", "_", colnames(lane_summary_df), fixed = TRUE)
colnames(lane_summary_df) <- gsub(")", "", colnames(lane_summary_df), fixed = TRUE)

top_unknown_barcodes_df <- read.csv(
  file.path(reports_dir, "Top_Unknown_Barcodes.csv"),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

undetermined_max_pcnt <- 10
quality_score_min <- 30
sample_reads_min <- 10
```

```{r lane_summary}
read_sum_df <- aggregate(PF_Clusters ~ Sample, data = lane_summary_df, FUN = sum)
undetermined_summary <- lane_summary_df[lane_summary_df[["Sample"]] == "Undetermined", ]
lane_sample_summary <- lane_summary_df[lane_summary_df[["Sample"]] != "Undetermined", ]

lane_reads_plot <- ggplot(lane_summary_df, aes(x = Sample, y = PF_Clusters / 1000000)) +
  geom_bar(aes(fill = Lane), position = "dodge", stat = "identity") +
  ylab("Millions of Reads") +
  xlab("Sample") +
  ggtitle("Barcode Matched Read Count per Sample per Lane") +
  coord_flip()

sample_reads_plot <- ggplot(read_sum_df, aes(x = Sample, y = PF_Clusters / 1000000)) +
  geom_bar(position = "dodge", stat = "identity") +
  ylab("Millions of Reads") +
  xlab("Sample") +
  ggtitle("Total Barcode Matched Read Count per Sample") +
  coord_flip() +
  geom_hline(yintercept = sample_reads_min, color = "red")

pcnt_undetermined_plot <- ggplot(undetermined_summary, aes(x = Lane, y = pcnt_of_thelane)) +
  geom_bar(position = "dodge", stat = "identity") +
  ggtitle("Percent Undetermined Reads per Lane") +
  ylab("Percent of Lane") +
  xlab("Lane") +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10)) +
  geom_hline(yintercept = undetermined_max_pcnt, color = "red")

pcnt_sample_plot <- ggplot(lane_sample_summary, aes(x = Sample, y = pcnt_of_thelane)) +
  geom_bar(aes(fill = Lane), position = "dodge", stat = "identity") +
  ggtitle("Percent Sample Reads per Lane") +
  ylab("Percent of Lane") +
  xlab("Lane") +
  coord_flip()

lane_quality_score_plot <- ggplot(lane_summary_df, aes(x = Sample, y = Mean_QualityScore)) +
  geom_bar(aes(fill = Lane), position = "dodge", stat = "identity") +
  ylab("Mean Quality Score") +
  xlab("Sample") +
  ggtitle("Mean Quality Score per Sample per Lane") +
  coord_flip() +
  geom_hline(yintercept = quality_score_min, color = "red")

pcnt_perfect_barcode_plot <- ggplot(lane_sample_summary, aes(x = Sample, y = pcnt_Perfectbarcode)) +
  geom_bar(aes(fill = Lane), position = "dodge", stat = "identity") +
  ylab("Percent Perfect Barcode") +
  xlab("Sample") +
  ggtitle("Percent of Reads with Perfect Barcodes per Sample per Lane") +
  coord_flip()
```

```{r sample_table}
run_project <- unique(as.character(lane_sample_summary[["Project"]]))
summary_cols <- c("Lane", "Project", "Sample", "Barcode_sequence")
run_lanes <- unique(as.character(lane_sample_summary[["Lane"]]))
run_summary <- list()
for (lane in run_lanes) {
  lane_df <- lane_sample_summary[which(as.character(lane_sample_summary[["Lane"]]) == lane), summary_cols]
  run_summary[[lane]] <- lane_df
}
print_lane_df <- function(df, name) {
  mycat(sprintf("### Lane %s \n\n", name))
  print(kable(df, row.names = FALSE))
}
```

```{r avg_undetermined_rating}
avg_undetermined <- mean(undetermined_summary[["pcnt_of_thelane"]])
print_undetermined_rating <- function(avg_undetermined) {
  determined_reads <- 100 - avg_undetermined
  if (determined_reads > 90) {
    print_big_color_boxed_text(message = "Undetermined Reads Score: A", border_color = "blue", text_color = "blue")
  } else if (determined_reads > 80) {
    print_big_color_boxed_text(message = "Undetermined Reads Score: B", border_color = "green", text_color = "green")
  } else if (determined_reads > 70) {
    print_big_color_boxed_text(message = "Undetermined Reads Score: C", border_color = "yellow", text_color = "yellow")
  } else if (determined_reads > 60) {
    print_big_color_boxed_text(message = "Undetermined Reads Score: D", border_color = "red", text_color = "red")
  } else {
    print_big_color_boxed_text(message = "Undetermined Reads Score: F", border_color = "maroon", text_color = "maroon")
  }
}
print_big_color_boxed_text <- function(message, border_color, text_color) {
  mycat(sprintf(
    '<p style="border:3px; border-style:solid; border-color:%s; padding: 1em;"><span style="color:%s"><big><big><big>%s</big></big></big></span></p>',
    border_color, text_color, message
  ))
}
```

# Summary

```{r, results='asis'}
if (nzchar(run_id)) {
  mycat(sprintf("**Run ID:** %s  \n\n", run_id))
}
```

Project:

```{r}
mycat(run_project)
```

## Samples {.tabset .tabset-pills}

```{r, results='asis'}
for (i in seq_along(names(run_summary))) {
  name <- names(run_summary)[i]
  df <- run_summary[i]
  print_lane_df(df = df, name = name)
}
```

## Sample Metrics

```{r}
print(sample_reads_plot)
```

- Each sample must have at least `r sample_reads_min` million reads to qualify for NGS580 panel analysis.

```{r, results='asis'}
kable(read_sum_df)
```

# Lane Metrics

## Undetermined Reads

```{r, results='asis'}
print_undetermined_rating(avg_undetermined = avg_undetermined)
```

```{r}
print(pcnt_undetermined_plot)
```

- Each lane must have less than `r undetermined_max_pcnt`% undetermined reads to qualify for NGS580 panel analysis.

## Read Qualities

```{r, fig.height=12}
print(lane_reads_plot)
print(pcnt_sample_plot)
print(pcnt_perfect_barcode_plot)
print(lane_quality_score_plot)
```

- Each sample & lane must have a mean quality score of at least `r quality_score_min` to qualify for NGS580 panel analysis.

# Top Unknown Barcodes

```{r, results='asis'}
kable(head(top_unknown_barcodes_df, 10), row.names = FALSE)
```

```{r, results='asis'}
if (nrow(top_unknown_barcodes_df) > 10) {
  mycat(sprintf("\n\n*(Showing first 10 of %d rows.)*\n", nrow(top_unknown_barcodes_df)))
}
```

# System Information

```{r}
mycat(sprintf("System:\n%s\n%s",
              tryCatch(system("hostname", intern = TRUE)[1], error = function(e) "NA"),
              tryCatch(system("uname -srv", intern = TRUE)[1], error = function(e) "NA")))
mycat(sprintf("System user:\n%s", tryCatch(system("whoami", intern = TRUE)[1], error = function(e) "NA")))
mycat(sprintf("System location:\n%s", tryCatch(getwd(), error = function(e) "NA")))
mycat(sprintf("BCL Convert reports directory:\n%s", normalizePath(reports_dir)))
mycat(sprintf("Time and Date of report creation:\n%s", format(Sys.time())))
sessionInfo()
```
